---
title: "Using the CWUKNN Package"
author: "Matthew C. Hansen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CWU KNN Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The `CWUKNN` package is a implementation of the KNN problem, solved with threading and GPU computing. This vignette provides a guide on how to get started using the package.

The package includes multiple modes of execution, which can be switched between depending on the user's preference:
  - C++ (For threaded CPU execution)
  - CUDA (For newer Nvidia GPU cards)
  - OpenCL (For most Graphics Cards)

## Purpose

# Installation

To install the package, use the following steps:
*If using OpenCL*:
  On Windows: Add OPENCL_HOME as a new System Environment Variable, linking to the path where your OpenCL installation is stored. If using the Nvidia GPU Computing Toolkit, for example, this should be `C:\Program Files\NVIDIA GPU Computing Toolkit\CUDA\v12.8`

Then, run:
```{r, eval=FALSE}
install.packages("CWUKNN")
```

The installation should build each piece of the program, and will different implementations based on what the user's machine supports (EX: With or without CUDA or OpenCL support).

To remove the package, simply run:
```{r, eval=FALSE}
install.packages("CWUKNN")
```

# KNN Function

`CWUKNN` implements a single function: `KNN(x, y, k, mode, precision)`, which calculates the Information Energy (IE) values between two rows, using various options on how the program executes its code.

Example execution:
```{r, eval=FALSE}
# Define K values to search for
k = 5000;

# Take in some data:
data <- fread("data.csv", skip = 2)
column_0 <- data$COLUMN0
column_1 <- data$COLUMN1

## Calculate the IE between column_0 and column_1

# Use C++ implementation, with double precision
result_cpp <- KNN(column_0, column_1, k, mode="", precision="double");
# Use CUDA implementation, with float precision
result_cuda <- KNN(column_0, column_1, k, mode="cuda", precision="float");
# Use OpenCL implementation, with float precision
result_opencl <- KNN(column_0, column_1, k, mode="opencl", precision="float");

# result_cpp, result_cuda, and result_opencl contain single IE values.
```